---
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
    # github_document:
    # html_preview: yes
    # toc: yes
    # toc_depth: 2
editor_options: 
  chunk_output_type: inline
---

Processing and plotting the analysis from the real EEG participant data set taken from study by Bieniek and colleagues (2016).

# Dependencies
```{r setup, message=FALSE}
library(ggplot2)
library(tidyverse)
library(tibble)
library(cowplot)
library(beepr)
library(Rfast)
library(changepoint)
library(mcp)
library(rjags)
library(mutoss)
library(ecp)
library(readr)
library(bcp)
source("./code/functions.R")
source("./code/theme_gar.txt")
# Edit `one_over_f` function from `primer` package to control variance (Stevens, 2009). 
# Original function is available on [GitHub](https://github.com/HankStevens/primer).
# Copyright Hank Stevens.
source("./code/one_over_f.R")
# Load template: true onset = 160 ms, F=81, max at F=126
source("./code/erp_template.R")
# R version of Matlab code from Yeung et al. 2004
source("./code/eeg_noise.R")
# Lodaing the code that gives median instead of mean values in mcp summary table
source("./code/mcp_median_function.R")
# to use with eeg_noise function
meanpower <- unlist(read.table("./code/meanpower.txt"))
s <- seq(-300,600,2) # time vector for Bieniek data (from -300 to 600 ms)

# x axis range for plots 
xa <- -150 # minimum 
xb <- 250 # maximum
```

## Example time course from participant 1
```{r message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
library(readr)
p1 <- read_csv("./ftonsets_results_p1_p120_1000_permutations/participant_p1_s1_maxt2.txt", col_names = NA)


df <- tibble(x = s,
             y = p1$X1)

p.mcp <- ggplot(df, aes(x, y)) + theme_gar + 
  geom_line(linewidth = 1) +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("participant 1 maxt2"))
```

## Example time course from participant 2
```{r eval=FALSE}
p2 <- read_csv("./ftonsets_results_p1_p120_1000_permutations/participant_p2_s1_maxt2.txt", col_names = NA)

df <- tibble(x = s,
             y = p2$X1)

p.mcp <- ggplot(df, aes(x, y)) + theme_gar + 
  geom_line(linewidth = 1) +
  # geom_vline(xintercept = true_onset) +
  # geom_vline(xintercept = summary_fit[1,2], linetype = "dotted") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("participant 2 maxt2"))
```

```{r eval=FALSE}
p2 <- read_csv("./ftonsets_results_p1_p120_1000_permutations/participant_p2_s2_maxt2.txt", col_names = NA)

options(mc.cores = 3)
df <- tibble(x = s, y = p2$X1)

model <- list(
  y ~ 1 + sigma(1),
    ~ 0 + x + sigma(1)
)

prior <- list(
  cp_1 = "dunif(50,150)"
)

fit <- mcp(model, prior = prior, data = df, cores = 3, chains = 3)
summary.fit.median <- summary.mcpfit.median(fit)
pt.mcp <- round(summary.fit.median[1,2],0)

p.mcp <- ggplot(df, aes(x, y)) + theme_gar + 
  geom_line(linewidth = 1) +
  # geom_vline(xintercept = true_onset) +
  geom_vline(xintercept = pt.mcp, linetype = "solid") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("mcp estimated onset =",pt.mcp,"ms"))
```

## Apply BinSeg
```{r eval=FALSE}
res.cp <- cpt.meanvar(p2$X1, method = "BinSeg", Q=2)
pt.cp <- s[res.cp@cpts[1]]

df <- tibble(x = s,
             y = p2$X1)

p.cp <- ggplot(df, aes(x, y)) + theme_gar + 
  geom_line(linewidth = 1) +
  # geom_vline(xintercept = true_onset) +
  geom_vline(xintercept = pt.cp, linetype = "solid") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("BinSeg estimated onset =",pt.cp,"ms"))
```


## Apply PELT
```{r eval=FALSE}

res.pelt <- cpt.meanvar(p2$X1, method = "PELT", penalty = "Manual", pen.value = 30*log(length(p2$X1)))
pt.pelt <- s[res.pelt@cpts[1]]

df <- tibble(x = s,
             y = p2$X1)

p.pelt <- ggplot(df, aes(x, y)) + theme_gar + 
  geom_line(linewidth = 1) +
  # geom_vline(xintercept = true_onset) +
  geom_vline(xintercept = pt.pelt, linetype = "solid") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("PELT estimated onset =",pt.pelt,"ms"))
```


## Apply cp3o_delta
```{r eval=FALSE}

p2.matrix <- matrix(p2$X1, ncol = 1)

res.ecp <- e.cp3o_delta(Z = p2.matrix, K = 7, alpha = 1)
pt.ecp <- get_earliest_cp(res.ecp$estimates, s)

df <- tibble(x = s,
             y = p2$X1)

p.ecp <- ggplot(df, aes(x, y)) + theme_gar + 
  geom_line(linewidth = 1) +
  # geom_vline(xintercept = true_onset) +
  geom_vline(xintercept = pt.ecp, linetype = "solid") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("cp3o estimated onset =",pt.ecp,"ms"))
```

## Apply mcp
```{r eval=FALSE}
options(mc.cores = 3)
df <- tibble(x = s, y = p2$X1)

model <- list(
  y ~ 1 + sigma(1),
    ~ 0 + x + sigma(1)
)

prior <- list(
  cp_1 = "dunif(50,150)"
)

fit <- mcp(model, prior = prior, data = df, cores = 3, chains = 3)
summary.fit.median <- summary.mcpfit.median(fit)
pt.mcp <- round(summary.fit.median[1,2],0)

p.mcp <- ggplot(df, aes(x, y)) + theme_gar + 
  geom_line(linewidth = 1) +
  # geom_vline(xintercept = true_onset) +
  geom_vline(xintercept = pt.mcp, linetype = "solid") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("mcp estimated onset =",pt.mcp,"ms"))
```

## Apply bcp
```{r eval=FALSE}

res <- bcp(df$y, p0 = 0.1, mcmc = 3000)
pt.bcp <- s[which.max(res$posterior.prob)]

p.bcp <- ggplot(df, aes(x, y)) + theme_gar + 
  geom_line(linewidth = 1) +
  # geom_vline(xintercept = true_onset) +
  geom_vline(xintercept = pt.bcp, linetype = "solid") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("bcp estimated onset =",pt.bcp,"ms"))
p.bcp
```


## Cluster based test results
```{r include=FALSE, message=FALSE, eval=FALSE}

# reading in cluster sum onsets from the file for session 1
cs.data.s1 <- read_csv("./ftonsets_results_p1_p120_1000_permutations/participants_s1_onsets.txt", col_names = FALSE)
cs.data.s1 <- as.numeric(unlist(cs.data.s1))
cs.data.s1 <- tibble(session1.cp = cs.data.s1)
row.names(cs.data.s1) <- paste("participant", seq_len(nrow(cs.data.s1)))

# reading in session 2
cs.data.s2 <- read_csv("./ftonsets_results_p1_p120_1000_permutations/participants_s2_onsets.txt", col_names = FALSE)
cs.data.s2 <- as.numeric(unlist(cs.data.s2))
cs.data.s2 <- tibble(session2.cp = cs.data.s2)
row.names(cs.data.s2) <- paste("participant", seq_len(nrow(cs.data.s2)))

temp.cs.data.s1 <- cs.data.s1
temp.cs.data.s1$participant.id <- row.names(temp.cs.data.s1)
temp.cs.data.s2 <- cs.data.s2
temp.cs.data.s2$participant.id <- row.names(temp.cs.data.s2)

cs.data <- merge(temp.cs.data.s1, temp.cs.data.s2, by = "participant.id")
row.names(cs.data) <- cs.data$participant.id
cs.data$participant.id <- NULL
rm(temp.cs.data.s1, temp.cs.data.s2)
cs.data <- cs.data[complete.cases(cs.data),]

# plotting graph showing participant 2 curve with an onset location estimated by cluster sum
cs.onset <- cs.data["participant 2",1]

df <- tibble(x = s,
             y = p2$X1)

p.cs <- ggplot(df, aes(x, y)) + theme_gar + 
  geom_line(linewidth = 1) +
  # geom_vline(xintercept = true_onset) +
  geom_vline(xintercept = cs.onset, linetype = "solid") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("cluster based correction onset =",cs.onset,"ms"))
p.cs
```


```{r fig.height=8, fig.width=10, eval=FALSE}
cowplot::plot_grid(p.cp, p.pelt, p.ecp, p.mcp,
                   nrow = 2,
                   labels = c("A", "B", "C", "D"),
                   label_size = 15)
```


## Participant 2 onset in session 2

```{r eval=FALSE}
p2 <- read_csv("./ftonsets_results_p1_p120_1000_permutations/participant_p2_s2_maxt2.txt", col_names = NA)

df <- tibble(x = s,
             y = p2$X1)

# No effect localisation algorithm
ggplot(df, aes(x, y)) + theme_gar + 
  geom_line(linewidth = 1) +
  # geom_vline(xintercept = true_onset) +
  # geom_vline(xintercept = summary_fit[1,2], linetype = "dotted") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("participant 2 maxt2"))


# BinSeg
res.cp <- cpt.meanvar(p2$X1, method = "BinSeg", Q=2)
pt.cp <- s[res.cp@cpts[1]]

df <- tibble(x = s,
             y = p2$X1)

p.cp <- ggplot(df, aes(x, y)) + theme_gar + 
  geom_line(linewidth = 1) +
  # geom_vline(xintercept = true_onset) +
  geom_vline(xintercept = pt.cp, linetype = "solid") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("BinSeg estimated onset =",pt.cp,"ms"))


# PELT
res.pelt <- cpt.meanvar(p2$X1, method = "PELT", penalty = "Manual", pen.value = 30*log(length(p2$X1)))
pt.pelt <- s[res.pelt@cpts[1]]

df <- tibble(x = s,
             y = p2$X1)

p.pelt <- ggplot(df, aes(x, y)) + theme_gar + 
  geom_line(linewidth = 1) +
  # geom_vline(xintercept = true_onset) +
  geom_vline(xintercept = pt.pelt, linetype = "solid") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("PELT estimated onset =",pt.pelt,"ms"))


# mcp
options(mc.cores = 3)
df <- tibble(x = s, y = p2$X1)

model <- list(
  y ~ 1 + sigma(1),
    ~ 0 + x + sigma(1)
)

prior <- list(
  cp_1 = "dunif(50,150)"
)

fit <- mcp(model, prior = prior, data = df, cores = 3, chains = 3)
summary.fit.median <- summary.mcpfit.median(fit)
pt.mcp <- round(summary.fit.median[1,2],0)

p.mcp <- ggplot(df, aes(x, y)) + theme_gar + 
  geom_line(linewidth = 1) +
  geom_vline(xintercept = pt.mcp, linetype = "solid") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("mcp estimated onset =",pt.mcp,"ms"))



# Function to restrict cp3o_delta to record the first detected change point in the 150-250ms time window
# get_earliest_cp <- function(cps, Xf) {
#   # Filter the change points based on the time window
#   valid_cps <- cps[Xf[cps] >= 150 & Xf[cps] <= 250]
# 
#   if (length(valid_cps) > 0) {
#     # Find the earliest point within the valid time window
#     earliest_index <- min(valid_cps)
#     # Convert to time course using the time vector
#     earliest_time <- Xf[earliest_index]
#     return(earliest_time)
#   } else {
#     return(NA)  # Return NA if no valid change points were found in the specified time window
#   }
# }



# cp3o
p2.matrix <- matrix(p2$X1, ncol = 1)

res.ecp <- e.cp3o_delta(Z = p2.matrix, K = 7, alpha = 1)
pt.ecp <- get_earliest_cp(res.ecp$estimates, s)
# pt.ecp <- s[res.ecp$estimates]

df <- tibble(x = s,
             y = p2$X1)

p.ecp <- ggplot(df, aes(x, y)) + theme_gar + 
  geom_line(linewidth = 1) +
  # geom_vline(xintercept = true_onset) +
  geom_vline(xintercept = pt.ecp, linetype = "solid") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("cp3o estimated onset =",pt.ecp,"ms"))

p.cp
p.pelt
p.ecp
p.mcp
```


## Calculating the difference in estimates between the methods across both testing sessions
### BinSeg
```{r include=FALSE, message=FALSE, warning=FALSE}
p2.s1 <- read_csv("./p1_p120_2_sessions/participant_p2_s1_maxt2.txt", col_names = NA)
p2.s2 <- read_csv("./p1_p120_2_sessions/participant_p2_s2_maxt2.txt", col_names = NA)

res.cp <- cpt.meanvar(p2.s1$X1, method = "BinSeg", Q=2)
pt1.cp <- s[res.cp@cpts[1]]

res.cp <- cpt.meanvar(p2.s1$X1, method = "BinSeg", Q=2)
pt2.cp <- s[res.cp@cpts[1]]

df <- tibble(x = s,
             y = p2.s1$X1,
             z = p2.s2$X1)

p.bs <- ggplot(df, aes(x)) + theme_gar + 
  geom_line(aes(y = y), colour = "#E69F00", linewidth = 1) +
  geom_line(aes(y = z), colour = "black", linewidth = 1) +
  geom_vline(xintercept = pt1.cp, linetype = "solid", colour = "#E69F00") +
  geom_vline(xintercept = pt2.cp, linetype = "solid", colour = "black") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("BinSeg: Session 1 =",pt1.cp,"ms, Session 2 =",pt2.cp,"ms")) + 
  theme(plot.title = element_text(size = 15))
```


## PELT
```{r inlcude=FALSE, message=FALSE, warning=FALSE}
res.pelt <- cpt.meanvar(p2.s1$X1, method = "PELT", penalty = "Manual", pen.value = 30*log(length(p2.s1$X1)))
pt1.pelt <- s[res.pelt@cpts[1]]

res.pelt <- cpt.meanvar(p2.s2$X1, method = "PELT", penalty = "Manual", pen.value = 30*log(length(p2.s2$X1)))
pt2.pelt <- s[res.pelt@cpts[1]]

df <- tibble(x = s,
             y = p2.s1$X1,
             z = p2.s2$X1)

p.pelt <- ggplot(df, aes(x)) + theme_gar + 
  geom_line(aes(y = y), colour = "#E69F00", linewidth = 1) +
  geom_line(aes(y = z), colour = "black", linewidth = 1) +
  geom_vline(xintercept = pt1.pelt, linetype = "solid", colour = "#E69F00") +
  geom_vline(xintercept = pt2.pelt, linetype = "solid", colour = "black") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("PELT: Session 1 =",pt1.pelt,"ms, Session 2 =",pt2.pelt,"ms")) + 
  theme(plot.title = element_text(size = 15))
```


### cp3o
```{r include=FALSE, warning=FALSE, message=FALSE}

p2.s1.matrix <- matrix(p2.s1$X1, ncol = 1)
p2.s2.matrix <- matrix(p2.s2$X1, ncol = 1)

res.ecp <- e.cp3o_delta(Z = p2.s1.matrix, K = 8, alpha = 1)
pt1.ecp <- get_earliest_cp(res.ecp$estimates, s)

res.ecp <- e.cp3o_delta(Z = p2.s2.matrix, K = 8, alpha = 1)
pt2.ecp <- get_earliest_cp(res.ecp$estimates, s)

df <- tibble(x = s,
             y = p2.s1$X1,
             z = p2.s2$X1)

p.cp3o <- ggplot(df, aes(x)) + theme_gar + 
  geom_line(aes(y = y), colour = "#E69F00", linewidth = 1) +
  geom_line(aes(y = z), colour = "black", linewidth = 1) +
  geom_vline(xintercept = pt1.ecp, linetype = "solid", colour = "#E69F00") +
  geom_vline(xintercept = pt2.ecp, linetype = "solid", colour = "black") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("cp3o_delta: Session 1 =",pt1.ecp,"ms, Session 2 =",pt2.ecp,"ms")) + 
  theme(plot.title = element_text(size = 15))
```

## mcp
```{r include=FALSE, message=FALSE, warning=FALSE}
options(mc.cores = 3)
df.1 <- tibble(x = s, y = p2.s1$X1)
df.2 <- tibble(x = s, y = p2.s2$X1)

model <- list(
  y ~ 1 + sigma(1),
    ~ 0 + x + sigma(1)
)

prior <- list(
  cp_1 = "dunif(50,150)"
)

fit1 <- mcp(model, prior = prior, data = df.1, cores = 3, chains = 3)
summary.fit.median1 <- summary.mcpfit.median(fit1)
pt1.mcp <- round(summary.fit.median1[1,2],0)

fit2 <- mcp(model, prior = prior, data = df.2, cores = 3, chains = 3)
summary.fit.median2 <- summary.mcpfit.median(fit2)
pt2.mcp <- round(summary.fit.median2[1,2],0)

df <- tibble(x = s,
             y = p2.s1$X1,
             z = p2.s2$X1)

p.mcp <- ggplot(df, aes(x)) + theme_gar + 
  geom_line(aes(y = y), colour = "#E69F00", linewidth = 1) +
  geom_line(aes(y = z), colour = "black", linewidth = 1) +
  geom_vline(xintercept = pt1.mcp, linetype = "solid", colour = "#E69F00") +
  geom_vline(xintercept = pt2.mcp, linetype = "solid", colour = "black") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("mcp: Session 1 =",pt1.mcp,"ms, Session 2 =",pt2.mcp,"ms")) + 
  theme(plot.title = element_text(size = 15))
```

## bcp
```{r include=FALSE, message=FALSE, warning=FALSE}
res1 <- bcp(df$y, p0 = 0.1, mcmc = 3000)
res2 <- bcp(df$z, p0 = 0.1, mcmc = 3000)

pt1.bcp <- s[which.max(res1$posterior.prob)]
pt2.bcp <- s[which.max(res$posterior.prob)]

p.bcp <- ggplot(df, aes(x)) + theme_gar + 
  geom_line(aes(y = y), colour = "#E69F00", linewidth = 1) +
  geom_line(aes(y = z), colour = "black", linewidth = 1) +
  geom_vline(xintercept = pt1.bcp, linetype = "solid", colour = "#E69F00") +
  geom_vline(xintercept = pt2.bcp, linetype = "solid", colour = "black") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("bcp: Session 1 =",pt1.bcp,"ms, Session 2 =",pt2.bcp,"ms")) + 
  theme(plot.title = element_text(size = 15))
```


## cluster-sum
```{r include=FALSE, warning=FALSE, message=FALSE}
cs.onset.p2.s1 <- cs.data["participant 30",1]
cs.onset.p2.s2 <- cs.data["participant 30",2]

df <- tibble(x = s,
             y = p2.s1$X1,
             z = p2.s2$X1)

p.cs <- ggplot(df, aes(x)) + theme_gar + 
  geom_line(aes(y = y), colour = "#E69F00", linewidth = 1) +
  geom_line(aes(y = z), colour = "black", linewidth = 1) +
  geom_vline(xintercept = cs.onset.p2.s1, linetype = "solid", colour = "#E69F00") +
  geom_vline(xintercept = cs.onset.p2.s2, linetype = "solid", colour = "black") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("CS: session 1 =",cs.onset.p2.s1,"ms, session 2 =",cs.onset.p2.s2,"ms")) + 
  theme(plot.title = element_text(size = 15))
```

```{r fig.height=8, fig.width=10}
cowplot::plot_grid(p.cs, p.bs, p.pelt, p.cp3o, p.mcp, p.bcp,
                   nrow = 3,
                   labels = c("A", "B", "C", "D", "E", "F"),
                   label_size = 15)

# ggsave(filename = "./figures/real_data_2session_demo.pdf", width = 12, height = 10)
```

## Now doing the same but across the 75 participants tested twice.
### Read in the data
```{r include=FALSE}
data.directory <- "./p1_p120_2_sessions"

participant.data <- list()

for (i in 1:90) {
  
  file.s1 <- sprintf("%s/participant_p%d_s1_maxt2.txt", data.directory, i)
  file.s2 <- sprintf("%s/participant_p%d_s2_maxt2.txt", data.directory, i)
  
  if (file.exists(file.s1) && file.exists(file.s2)) {
    
    data.s1 <- read_csv(file.s1, col_names = FALSE)
    data.s2 <- read_csv(file.s2, col_names = FALSE)
    
    combined.data <- data.frame(session1 = data.s1$X1, session2 = data.s2$X1)

    participant.data[[paste("participant", i)]] <- combined.data
  }
}
```

## Apply methods
### BinSeg
```{r eval=FALSE, warning=FALSE}

change.points.cp <- list()


for (participant_id in names(participant.data)) {

  participant.df <- participant.data[[participant_id]]

  res.cp1 <- cpt.meanvar(participant.df$session1, method = "BinSeg", Q=2)
  pt1.cp <- s[res.cp1@cpts[1]]

  res.cp2 <- cpt.meanvar(participant.df$session2, method = "BinSeg", Q=2)
  pt2.cp <- s[res.cp2@cpts[1]]

  change.points.cp[[participant_id]] <- list(session1.cp = pt1.cp, session2.cp = pt2.cp)
}

change.points.cp.df <- do.call(rbind, lapply(change.points.cp, function(x) as.data.frame(t(unlist(x)))))
rownames(change.points.cp.df) <- names(change.points.cp)
save(change.points.cp, file = "./data/change.points.cp")
save(change.points.cp.df, file = "./data/change.points.cp.df")
```

### PELT
```{r eval=FALSE}
change.points.pelt <- list()

for (participant_id in names(participant.data)) {
  
  participant.df <- participant.data[[participant_id]]
  
  res.pelt1 <- cpt.meanvar(participant.df$session1, method = "PELT", penalty = "Manual", pen.value = 30*log(length(p2$X1)))
  pt1.pelt <- s[res.pelt1@cpts[1]]
  
  res.pelt2 <- cpt.meanvar(participant.df$session2, method = "PELT", penalty = "Manual", pen.value = 30*log(length(p2$X1)))
  pt2.pelt <- s[res.pelt2@cpts[1]]
  
  change.points.pelt[[participant_id]] <- list(session1.cp = pt1.pelt, session2.cp = pt2.pelt)
}

change.points.pelt.df <- do.call(rbind, lapply(change.points.pelt, function(x) as.data.frame(t(unlist(x)))))
rownames(change.points.pelt.df) <- names(change.points.cp)
save(change.points.pelt, file = "./data/change.points.pelt")
save(change.points.pelt.df, file = "./data/change.points.pelt.df")
```

### ecp
```{r eval=FALSE}
change.points.ecp <- list()

for (participant_id in names(participant.data)) {
  
  participant.df <- participant.data[[participant_id]]
  
  session1.matrix <- matrix(participant.df$session1, ncol = 1)
  session2.matrix <- matrix(participant.df$session2, ncol = 1)
  
  res.ecp1 <- e.cp3o_delta(Z = session1.matrix, K = 8, alpha = 1)
  pt1.ecp <- get_earliest_cp(res.ecp1$estimates, s)
  
  res.ecp2 <- e.cp3o_delta(Z = session2.matrix, K = 8, alpha = 1)
  pt2.ecp <- get_earliest_cp(res.ecp2$estimates, s)
  
  change.points.ecp[[participant_id]] <- list(session1.cp = pt1.ecp, session2.cp = pt2.ecp)
}

change.points.ecp.df <- do.call(rbind, lapply(change.points.ecp, function(x) as.data.frame(t(unlist(x)))))
rownames(change.points.ecp.df) <- names(change.points.ecp)

save(change.points.ecp, file = "./data/change.points.ecp")
save(change.points.ecp.df, file = "./data/change.points.ecp.df")
```

### mcp
```{r message=FALSE, include=FALSE, eval=FALSE}
options(mc.cores = 3)
change.points.mcp <- list()

model <- list(
  y ~ 1 + sigma(1),
    ~ 0 + x + sigma(1)
)

prior <- list(
  cp_1 = "dunif(50, 150)"
)

for (participant_id in names(participant.data)) {
  
  participant.df <- participant.data[[participant_id]]
  
  df.s1 <- tibble(x = s, y = participant.df$session1)
  df.s2 <- tibble(x = s, y = participant.df$session2)
  
  fit1 <- mcp(model, prior = prior, data = df.s1, cores = 3, chains = 3)
  summary.fit.median1 <- summary.mcpfit.median(fit1)
  pt1.mcp <- round(summary.fit.median1[1,2],0)
  
  fit2 <- mcp(model, prior = prior, data = df.s2, cores = 3, chains = 3)
  summary.fit.median2 <- summary.mcpfit.median(fit2)
  pt2.mcp <- round(summary.fit.median2[1,2],0)
  
  change.points.mcp[[participant_id]] <- list(session1.mcp = pt1.mcp, session2.mcp = pt2.mcp)
}

change.points.mcp.df <- do.call(rbind, lapply(change.points.mcp, function(x) as.data.frame(t(unlist(x)))))
rownames(change.points.mcp.df) <- names(change.points.mcp)
save(change.points.mcp, file = "./data/change.points.mcp")
save(change.points.mcp.df, file = "./data/change.points.mcp.df")
```

### bcp
```{r}
change.points.bcp <- list()

for (participant_id in names(participant.data)) {
  
  participant.df <- participant.data[[participant_id]]
  
  df.s1 <- tibble(x = s, y = participant.df$session1)
  df.s2 <- tibble(x = s, y = participant.df$session2)
  
  res1 <- bcp(df.s1$y, p0 = 0.1, mcmc = 3000)
  pt1.bcp <- s[which.max(res1$posterior.prob)]
  
  res2 <- bcp(df.s2$y, p0 = 0.1, mcmc = 3000)
  pt2.bcp <- s[which.max(res2$posterior.prob)]

  change.points.bcp[[participant_id]] <- list(session1.bcp = pt1.bcp, session2.bcp = pt2.bcp)
}

change.points.bcp.df <- do.call(rbind, lapply(change.points.bcp, function(x) as.data.frame(t(unlist(x)))))
rownames(change.points.bcp.df) <- names(change.points.bcp)
save(change.points.bcp, file = "./data/change.points.bcp")
save(change.points.bcp.df, file = "./data/change.points.bcp.df")
```


#### Calculating mean and median distributions of onset for each method 
```{r}

load(file = "./data/change.points.cp.df")
load(file = "./data/change.points.mcp.df")
load(file = "./data/change.points.ecp.df")
load(file = "./data/change.points.pelt.df")
load(file = "./data/change.points.bcp.df")

###############################################
# median of differences between groups ----
session1.median.cp <- median(change.points.cp.df$session1.cp)
session2.median.cp <- median(change.points.cp.df$session2.cp)
change.points.cp.df$difference <- abs(change.points.cp.df$session1.cp - change.points.cp.df$session2.cp)
diff.median.cp <- median(change.points.cp.df$difference)

session1.median.pelt <- median(change.points.pelt.df$session1.cp)
session2.median.pelt <- median(change.points.pelt.df$session2.cp)
change.points.pelt.df$difference <- abs(change.points.pelt.df$session1.cp - change.points.pelt.df$session2.cp)
diff.median.pelt <- median(change.points.pelt.df$difference)

session1.median.ecp <- median(change.points.ecp.df$session1.cp, na.rm = TRUE)
session2.median.ecp <- median(change.points.ecp.df$session2.cp, na.rm = TRUE)
change.points.ecp.df$difference <- abs(change.points.ecp.df$session1.cp - change.points.ecp.df$session2.cp)
diff.median.ecp <- median(change.points.ecp.df$difference, na.rm = TRUE)

session1.median.mcp <- median(change.points.mcp.df$session1.mcp)
session2.median.mcp <- median(change.points.mcp.df$session2.mcp)
change.points.mcp.df$difference <- abs(change.points.mcp.df$session1.mcp - change.points.mcp.df$session2.mcp)
diff.median.mcp <- median(change.points.mcp.df$difference)

session1.median.bcp <- median(change.points.bcp.df$session1.bcp)
session2.median.bcp <- median(change.points.bcp.df$session2.bcp)
change.points.bcp.df$difference <- abs(change.points.bcp.df$session1.bcp - change.points.bcp.df$session2.bcp)
diff.median.bcp <- median(change.points.bcp.df$difference)

session1.median.cs <- median(cs.data$session1.cp)
session2.median.cs <- median(cs.data$session2.cp)
cs.data$difference <- abs(cs.data$session1.cp - cs.data$session2.cp)
diff.median.cs <- median(cs.data$difference)

print("Medians:")
print(paste("CS =",diff.median.cs))
print(paste("BinSeg =",diff.median.cp))
print(paste("mcp =",diff.median.mcp))
print(paste("PELT =",diff.median.pelt))
print(paste("cp3o_delta =",diff.median.ecp))
print(paste("bcp =",diff.median.bcp))

##############################################
# variances ----
print("Variance Session 1:")
print(paste("CS =",round(var(cs.data$session1.cp), digits=0)))
print(paste("BinSeg =",round(var(change.points.cp.df$session1.cp), digits=0)))
print(paste("mcp =",round(var(change.points.mcp.df$session1.mcp), digits=0)))
print(paste("PELT =",round(var(change.points.pelt.df$session1.cp), digits=0)))
print(paste("cp3o_delta =",round(var(change.points.ecp.df$session1.cp, na.rm = TRUE), digits=0)))
print(paste("bcp =",round(var(change.points.bcp.df$session1.bcp), digits=0)))

print("Variance Session 2:")
print(paste("CS =",round(var(cs.data$session2.cp), digits=0)))
print(paste("BinSeg =",round(var(change.points.cp.df$session2.cp), digits=0)))
print(paste("mcp =",round(var(change.points.mcp.df$session2.mcp), digits=0)))
print(paste("PELT =",round(var(change.points.pelt.df$session2.cp), digits=0)))
print(paste("cp3o_delta =",round(var(change.points.ecp.df$session2.cp, na.rm = TRUE), digits=0)))
print(paste("bcp =",round(var(change.points.bcp.df$session2.bcp), digits=0)))

print("Mean Variance:")
print(paste("CS =",round((var(cs.data$session1.cp)+var(cs.data$session2.cp))/2, digits=0)))
print(paste("BinSeg =",round((var(change.points.cp.df$session1.cp)+var(change.points.cp.df$session2.cp))/2, digits=0)))
print(paste("mcp =",round((var(change.points.mcp.df$session1.mcp)+var(change.points.mcp.df$session2.mcp))/2, digits=0)))
print(paste("PELT =",round((var(change.points.pelt.df$session1.cp)+var(change.points.pelt.df$session2.cp))/2, digits=0)))
print(paste("cp3o_delta =",round((var(change.points.ecp.df$session1.cp)+var(change.points.ecp.df$session2.cp))/2, digits=0)))
print(paste("bcp =",round((var(change.points.bcp.df$session1.bcp)+var(change.points.bcp.df$session2.bcp))/2, digits=0)))
```


## Plotting onset distributions 
#### BinSeg
```{r}
library(tidyr)

load(file = "./data/change.points.cp.df")

df <- change.points.cp.df |> pivot_longer(cols = everything(), names_to = "session", values_to = "change.points")

categ.palette <- c("#000000", "#E69F00", "#009E73", "#0072B2", "#D55E00", "#CC79A7")

p.cp1 <- ggplot(data = df, aes(x = change.points, color = session)) + theme_gar +
  geom_freqpoly(binwidth = 3, size = .7, na.rm = TRUE) +  
  scale_color_manual(values = categ.palette) +
  labs(x = "Change Point Onsets in ms", y = "Frequency", color = "Session") +
  geom_vline(xintercept = session1.median.cp, linetype = "solid", colour = "#000000") +
  geom_vline(xintercept = session2.median.cp, linetype = "solid", colour = "#E69F00") +
  ggtitle("B: BinSeg") +
  xlim(xa,xb) + ylim(0,15) + 
  theme(legend.position = "none")

df <- change.points.cp.df %>%
  mutate(participant_id = rownames(.))

p.cp2 <- ggplot(data = df, aes(x = session1.cp, y = session2.cp)) +
  geom_point(color = "#000000", size = 3, alpha = 0.6) +  
  geom_smooth(method = "lm", se = FALSE, color = "#E69F00", fullrange = TRUE) + 
  labs(x = "Session 1 Onset (ms)", y = "Session 2 Onset (ms)") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "#000000") +
  ggtitle("B: BinSeg") +
  xlim(xa, xb) + ylim(xa, xb) + 
  theme_gar +
  theme(legend.position = "none") +
  coord_fixed(ratio = 1)
```

### PELT
```{r}
load(file = "./data/change.points.pelt.df")
df <- change.points.pelt.df |> pivot_longer(cols = everything(), names_to = "session", values_to = "change.points")

categ.palette <- c("#000000", "#E69F00", "#009E73", "#0072B2", "#D55E00", "#CC79A7")

p.pelt1 <- ggplot(data = df, aes(x = change.points, color = session)) + theme_gar +
  geom_freqpoly(binwidth = 3, size = .7, na.rm = TRUE) +  
  scale_color_manual(values = categ.palette) +
  labs(x = "Change Point Onsets in ms", y = "Frequency", color = "Session") +
  geom_vline(xintercept = session1.median.pelt, linetype = "solid", colour = "#000000") +
  geom_vline(xintercept = session2.median.pelt, linetype = "solid", colour = "#E69F00") +
  ggtitle("C: PELT") + 
  xlim(xa,xb) + ylim(0,15) + 
  theme(legend.position = "none")

df <- change.points.pelt.df %>%
  mutate(participant_id = rownames(.))

p.pelt2 <- ggplot(data = df, aes(x = session1.cp, y = session2.cp)) +
  geom_point(color = "#000000", size = 3, alpha = 0.6) +  
  geom_smooth(method = "lm", se = FALSE, color = "#E69F00", fullrange = TRUE) + 
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "#000000") +
  labs(x = "Session 1 Onset (ms)", y = "Session 2 Onset (ms)") +
  ggtitle("C: PELT") +
  xlim(xa, xb) + ylim(xa, xb) + 
  theme_gar +
  theme(legend.position = "none") +
  coord_fixed(ratio = 1)
```

### ecp
```{r}
load(file = "./data/change.points.ecp.df")
df <- change.points.ecp.df |> pivot_longer(cols = everything(), names_to = "session", values_to = "change.points")

categ.palette <- c("#000000", "#E69F00", "#009E73", "#0072B2", "#D55E00", "#CC79A7")

p.ecp1 <- ggplot(data = df, aes(x = change.points, color = session)) + theme_gar +
  geom_freqpoly(binwidth = 3, size = .7, na.rm = TRUE) +  
  scale_color_manual(values = categ.palette) +
  labs(x = "Change Point Onsets in ms", y = "Frequency", color = "Session") +
  geom_vline(xintercept = session1.median.ecp, linetype = "solid", colour = "#000000") +
  geom_vline(xintercept = session2.median.ecp, linetype = "solid", colour = "#E69F00") +
  ggtitle("D: cp3o_delta") + 
  xlim(xa,xb) + ylim(0,15) + 
  theme(legend.position = "none")

df <- change.points.ecp.df %>%
  mutate(participant_id = rownames(.))

p.ecp2 <- ggplot(data = df, aes(x = session1.cp, y = session2.cp)) +
  geom_point(color = "#000000", size = 3, alpha = 0.6) +  
  geom_smooth(method = "lm", se = FALSE, color = "#E69F00", fullrange = TRUE) + 
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "#000000") +
  labs(x = "Session 1 Onset (ms)", y = "Session 2 Onset (ms)") +
  ggtitle("D: cp3o_delta") +
  xlim(xa, xb) + ylim(xa, xb) + 
  theme_gar +
  theme(legend.position = "none") +
  coord_fixed(ratio = 1)
```

### mcp
```{r}
load(file = "./data/change.points.mcp.df")
df <- change.point <- change.points.mcp.df |> pivot_longer(cols = everything(), names_to = "session", values_to = "change.points")

categ.palette <- c("#000000", "#E69F00", "#009E73", "#0072B2", "#D55E00", "#CC79A7")

p.mcp1 <- ggplot(data = df, aes(x = change.points, color = session)) + theme_gar +
  geom_freqpoly(binwidth = 3, size = .7, na.rm = TRUE) +  
  scale_color_manual(values = categ.palette) +
  labs(x = "Change Point Onsets in ms", y = "Frequency", color = "Session") +
  geom_vline(xintercept = session1.median.mcp, linetype = "solid", colour = "#000000") +
  geom_vline(xintercept = session2.median.mcp, linetype = "solid", colour = "#E69F00") +
  ggtitle("E: mcp") + 
  xlim(xa,xb) + ylim(0,15) +
  theme(legend.position = "none")

df <- change.points.mcp.df %>%
  mutate(participant_id = rownames(.))

p.mcp2 <- ggplot(data = df, aes(x = session1.mcp, y = session2.mcp)) +
  geom_point(color = "#000000", size = 3, alpha = 0.6) +  
  geom_smooth(method = "lm", se = FALSE, color = "#E69F00", fullrange = TRUE) + 
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "#000000") +
  labs(x = "Session 1 Onset (ms)", y = "Session 2 Onset (ms)") +
  ggtitle("E: mcp") +
  xlim(xa, xb) + ylim(xa, xb) + 
  theme_gar +
  theme(legend.position = "none") +
  coord_fixed(ratio = 1)
```

#### bcp
```{r}
load(file = "./data/change.points.bcp.df")
df <- change.points.bcp.df |> pivot_longer(cols = everything(), names_to = "session", values_to = "change.points")

categ.palette <- c("#000000", "#E69F00", "#009E73", "#0072B2", "#D55E00", "#CC79A7")

p.bcp1 <- ggplot(data = df, aes(x = change.points, color = session)) + theme_gar +
  geom_freqpoly(binwidth = 3, size = .7, na.rm = TRUE) +  
  scale_color_manual(values = categ.palette) +
  labs(x = "Change Point Onsets in ms", y = "Frequency", color = "Session") +
  geom_vline(xintercept = session1.median.bcp, linetype = "solid", colour = "#000000") +
  geom_vline(xintercept = session2.median.bcp, linetype = "solid", colour = "#E69F00") +
  ggtitle("E: bcp") + 
  xlim(xa,xb) + ylim(0,15) + 
  theme(legend.position = "none")

df <- change.points.bcp.df %>%
  mutate(participant_id = rownames(.))

p.ecp2 <- ggplot(data = df, aes(x = session1.bcp, y = session2.bcp)) +
  geom_point(color = "#000000", size = 3, alpha = 0.6) +  
  geom_smooth(method = "lm", se = FALSE, color = "#E69F00", fullrange = TRUE) + 
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "#000000") +
  labs(x = "Session 1 Onset (ms)", y = "Session 2 Onset (ms)") +
  ggtitle("E: bcp") +
  xlim(xa, xb) + ylim(xa, xb) + 
  theme_gar +
  theme(legend.position = "none") +
  coord_fixed(ratio = 1)
```

#### cluster sum
```{r}
df <- cs.data |> pivot_longer(cols = everything(), names_to = "session", values_to = "change.points")

categ.palette <- c("#000000", "#E69F00", "#009E73", "#0072B2", "#D55E00", "#CC79A7")

p.cs1 <- ggplot(data = df, aes(x = change.points, color = session)) + theme_gar +
  geom_freqpoly(binwidth = 3, size = .7, na.rm = TRUE) +  
  scale_color_manual(values = categ.palette) +
  labs(x = "Change Point Onsets in ms", y = "Frequency", color = "Session") +
  ggtitle("A: CS") + 
  geom_vline(xintercept = median(cs.data$session1.cp), linetype = "solid", colour = "#000000") +
  geom_vline(xintercept = median(cs.data$session2.cp), linetype = "solid", colour = "#E69F00") +
  xlim(xa,xb) + ylim(0,15) +
  theme(legend.position = "none")

df <- cs.data %>%
  mutate(participant_id = rownames(.))

p.cs2 <- ggplot(data = df, aes(x = session1.cp, y = session2.cp)) +
  geom_point(color = "#000000", size = 3, alpha = 0.6) +  
  geom_smooth(method = "lm", se = FALSE, color = "#E69F00", fullrange = TRUE) + 
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "#000000") +
  labs(x = "Session 1 Onset (ms)", y = "Session 2 Onset (ms)") +
  ggtitle("A: Cluster-based approach") +
  xlim(xa, xb) + ylim(xa, xb) + 
  theme_gar +
  theme(legend.position = "none") +
  coord_fixed(ratio = 1)
```

#### Joined figure 
```{r fig.height=6, fig.width=10}

joined.cs <- cs.data
joined.cs$difference <- NULL
colnames(joined.cs)[1] <- "session1.cs"
colnames(joined.cs)[2] <- "session2.cs"

joined.cp <- change.points.cp.df
joined.cp$difference <- NULL
colnames(joined.cp)[1] <- "session1.cp"
colnames(joined.cp)[2] <- "session2.cp"

joined.pelt <- change.points.pelt.df
joined.pelt$difference <- NULL
colnames(joined.pelt)[1] <- "session1.pelt"
colnames(joined.pelt)[2] <- "session2.pelt"

joined.ecp <- change.points.ecp.df
joined.ecp$difference <- NULL
colnames(joined.ecp)[1] <- "session1.ecp"
colnames(joined.ecp)[2] <- "session2.ecp"

joined.mcp <- change.points.mcp.df
joined.mcp$difference <- NULL
colnames(joined.mcp)[1] <- "session1.mcp"
colnames(joined.mcp)[2] <- "session2.mcp"

joined.bcp <- change.points.bcp.df
joined.bcp$difference <- NULL
colnames(joined.bcp)[1] <- "session1.bcp"
colnames(joined.bcp)[2] <- "session2.bcp"

joined.s1.s2 <- cbind(joined.cs, joined.cp, joined.pelt, joined.ecp, joined.mcp, joined.bcp)

point.size <- 1.3
point.alpha <- 0.5
line.width <- 0.7

p.s1.s2.2 <- ggplot() +
  geom_point(data = joined.s1.s2, aes(x = session1.cs, y = session2.cs, color = "CS"), size = point.size, alpha = point.alpha) +
  geom_smooth(data = joined.s1.s2, aes(x = session1.cs, y = session2.cs, color = "CS"), method = "lm", se = FALSE, fullrange = TRUE, linewidth = line.width) +
  geom_point(data = joined.s1.s2, aes(x = session1.cp, y = session2.cp, color = "BinSeg"), size = point.size, alpha = point.alpha) +
  geom_smooth(data = joined.s1.s2, aes(x = session1.cp, y = session2.cp, color = "BinSeg"), method = "lm", se = FALSE, fullrange = TRUE, linewidth = line.width) +
  geom_point(data = joined.s1.s2, aes(x = session1.ecp, y = session2.ecp, color = "cp3o_delta"), size = point.size, alpha = point.alpha) +
  geom_smooth(data = joined.s1.s2, aes(x = session1.ecp, y = session2.ecp, color = "cp3o_delta"), method = "lm", se = FALSE, fullrange = TRUE, linewidth = line.width) +
  geom_point(data = joined.s1.s2, aes(x = session1.pelt, y = session2.pelt, color = "PELT"), size = point.size, alpha = point.alpha) +
  geom_smooth(data = joined.s1.s2, aes(x = session1.pelt, y = session2.pelt, color = "PELT"), method = "lm", se = FALSE, fullrange = TRUE, linewidth = line.width) +
  geom_point(data = joined.s1.s2, aes(x = session1.mcp, y = session2.mcp, color = "mcp"), size = point.size, alpha = point.alpha) +
  geom_smooth(data = joined.s1.s2, aes(x = session1.mcp, y = session2.mcp, color = "mcp"), method = "lm", se = FALSE, fullrange = TRUE, linewidth = line.width) +
  geom_point(data = joined.s1.s2, aes(x = session1.bcp, y = session2.bcp, color = "bcp"), size = point.size, alpha = point.alpha) +
  geom_smooth(data = joined.s1.s2, aes(x = session1.bcp, y = session2.bcp, color = "bcp"), method = "lm", se = FALSE, fullrange = TRUE, linewidth = line.width) +
  labs(x = "Session 1 Onset (ms)", y = "Session 2 Onset (ms)", color = "Method") +
  xlim(-250, 600) + ylim(-250, 600) + 
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", color = "#000000") +
  ggtitle("A") +
  theme_gar +
  theme(legend.position = "bottom", legend.text = element_text(size = 12), legend.title = element_text(size = 13)) +
  coord_fixed(ratio = 1)

# Define color scale manually
p.s1.s2.2 <- p.s1.s2.2 + scale_color_manual(values = c(
  "CS" = "#0072B2",
  "BinSeg" = "#E69F00",
  "cp3o_delta" = "#009E73",
  "PELT" = "#D55E00",
  "mcp" = "#CC79A7",
  "bcp" = "#000000"
))

p.s1.s2.2 <- p.s1.s2.2 +
  annotate("rect", xmin = -50, xmax = 200, ymin = -50, ymax = 200, 
           color = "red", fill = NA, linetype = "dashed")

p.s1.s2.2.zoom <- ggplot() +
  geom_point(data = joined.s1.s2, aes(x = session1.cs, y = session2.cs, color = "CS"), size = point.size, alpha = point.alpha) +
  geom_point(data = joined.s1.s2, aes(x = session1.cp, y = session2.cp, color = "BinSeg"), size = point.size, alpha = point.alpha) +
  geom_point(data = joined.s1.s2, aes(x = session1.ecp, y = session2.ecp, color = "cp3o_delta"), size = point.size, alpha = point.alpha) +
  geom_point(data = joined.s1.s2, aes(x = session1.pelt, y = session2.pelt, color = "PELT"), size = point.size, alpha = point.alpha) +
  geom_point(data = joined.s1.s2, aes(x = session1.mcp, y = session2.mcp, color = "mcp"), size = point.size, alpha = point.alpha) +
  geom_point(data = joined.s1.s2, aes(x = session1.bcp, y = session2.bcp, color = "bcp"), size = point.size, alpha = point.alpha) +
  xlim(-50, 200) + ylim(-50, 200) + 
  xlab("Session 1 Onset (ms)") + ylab("Session 2 Onset (ms)") +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", color = "#000000") +
  theme_gar +
  theme(legend.position = "none") +
  ggtitle("B") +
  coord_fixed(ratio = 1)

p.s1.s2.2.zoom <- p.s1.s2.2.zoom + scale_color_manual(values = c(
  "CS" = "#0072B2", 
  "BinSeg" = "#E69F00",
  "cp3o_delta" = "#009E73",
  "PELT" = "#D55E00",
  "mcp" = "#CC79A7",
  "bcp" = "#000000"
), guide = FALSE)

library(patchwork)
fig <- p.s1.s2.2 + p.s1.s2.2.zoom + plot_layout(guides = 'collect')
fig <- fig & theme(legend.position = "bottom")
fig


# ggsave(filename = "./figures/test-retest_scatter.pdf", width = 8, height = 5)
```

```{r fig.width=8, fig.height=6}

data.long <- pivot_longer(
  joined.s1.s2,
  cols = everything(),
  names_to = c("method", "session"),
  names_sep = "\\.",
  values_to = "onset"
)

data.long <- data.long |> 
  mutate(across(.cols = everything(), .fns = ~ ifelse(.x < 0, NA, .x)))

data.long <- data.long |> 
  mutate(group = interaction(method, session))

# Calculate medians
medians <- data.long |> 
  group_by(group) |> 
  summarise(median = median(onset, na.rm = TRUE), .groups = 'drop') |> 
  mutate(group = factor(group, levels = c(
    "session1.cs", "session2.cs",
    "session1.cp", "session2.cp",
    "session1.mcp", "session2.mcp",
    "session1.pelt", "session2.pelt",
    "session1.ecp", "session2.ecp",
    "session1.bcp", "session2.bcp"
  )))


data.long <- data.long |> 
  mutate(group = factor(group, levels = c(
    "session1.ecp", "session2.ecp",
    "session1.pelt", "session2.pelt",
    "session1.mcp", "session2.mcp",
    "session1.cp", "session2.cp",
    "session1.cs", "session2.cs",
    "session1.bcp", "session2.bcp"
  ))) |> 
  mutate(group = fct_recode(group,
    "cp3o_delta Session 2" = "session1.ecp",
    "cp3o_delta Session 1" = "session2.ecp",
    "PELT Session 2" = "session1.pelt",
    "PELT Session 1" = "session2.pelt",
    "mcp Session 2" = "session1.mcp",
    "mcp Session 1" = "session2.mcp",
    "BinSeg Session 2" = "session1.cp",
    "BinSeg Session 1" = "session2.cp",
    "CS Session 2" = "session1.cs",
    "CS Session 1" = "session2.cs",
    "bcp Session 2" = "session1.bcp",
    "bcp Session 1" = "session2.bcp"))

# Custom colour mapping
color_map <- c(
  "CS Session 1" = "#0072B2", "CS Session 2" = "#0072B2",
  "BinSeg Session 1" = "#E69F00", "BinSeg Session 2" = "#E69F00",
  "mcp Session 1" = "#D55E00", "mcp Session 2" = "#D55E00",
  "PELT Session 1" = "#CC79A7", "PELT Session 2" = "#CC79A7",
  "cp3o_delta Session 1" = "#009E73", "cp3o_delta Session 2" = "#009E73",
  "bcp Session 1" = "grey10", "bcp Session 2"= "grey10"
)

p.box <- ggplot(data.long, aes(x = group, y = onset, fill = group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(width = 0.15, size = 1, alpha = 0.4, color = "black") +
  scale_fill_manual(values = color_map) +
  labs(x = "Method and Session", y = "Onset (ms)") +
  theme(legend.position = "none", plot.title = element_text(size = 20)) +
  theme_gar +
  ylim(0, 200) + 
  scale_fill_manual(values = color_map, guide = FALSE) +
  coord_flip()
p.box

ggsave(filename = "./figures/test-retest_box.pdf", width = 8, height = 5)
```


```{r fig.height=8, fig.width=10}
cowplot::plot_grid(p.cp1, p.pelt1, p.ecp1, p.mcp1, p.cs1,
                   nrow = 3,
                   labels = c("A", "B", "C", "D", "E"),
                   label_size = 15)
```

```{r fig.height=8, fig.width=8}

cowplot::plot_grid(p.cp2, p.pelt2, p.ecp2, p.mcp2, p.cs2,
                   nrow = 3,
                   labels = c("A", "B", "C", "D", "E"),
                   label_size = 15)
```

## Onset distributions for all participants from session 1
```{r include=FALSE}
data.directory <- "./p1_p120_session_1"

participant.data.s1 <- list()

for (i in 1:120) {
  
  file.s1 <- sprintf("%s/participant_p%d_s1_maxt2.txt", data.directory, i)
  
  if (file.exists(file.s1)) {
    
    data.s1 <- read_csv(file.s1, col_names = FALSE)

    participant.data.s1[[paste("participant", i)]] <- data.frame(session1 = data.s1$X1)
  }
}
```

### Identifying onsets
#### BinSeg
```{r eval=FALSE, warning=FALSE}
change.points.cp.s1 <- list()

for (participant_id in names(participant.data.s1)) {

  participant.df <- participant.data.s1[[participant_id]]

  res.cp <- cpt.meanvar(participant.df$session1, method = "BinSeg", Q=2)
  pt.cp <- s[res.cp@cpts[1]]


  change.points.cp.s1[[participant_id]] <- list(session1.cp = pt.cp)
}

change.points.cp.df.s1 <- do.call(rbind, lapply(change.points.cp.s1, function(x) as.data.frame(t(unlist(x)))))
rownames(change.points.cp.df.s1) <- names(change.points.cp.s1)
save(change.points.cp.df.s1, file = "./data/change.points.cp.df.s1")
```

#### PELT
```{r eval=FALSE}
change.points.pelt.s1 <- list()

for (participant_id in names(participant.data.s1)) {

  participant.df <- participant.data.s1[[participant_id]]

  res.pelt <- cpt.meanvar(participant.df$session1, method = "PELT", penalty = "Manual", pen.value = 30*log(length(p2$X1)))
  
  pt.pelt <- s[res.pelt@cpts[1]]


  change.points.pelt.s1[[participant_id]] <- list(session1.cp = pt.pelt)
}

change.points.pelt.df.s1 <- do.call(rbind, lapply(change.points.pelt.s1, function(x) as.data.frame(t(unlist(x)))))
rownames(change.points.pelt.df.s1) <- names(change.points.pelt.s1)
save(change.points.pelt.df.s1, file = "./data/change.points.pelt.df.s1")
```

#### ecp
```{r eval=FALSE}
change.points.ecp.s1 <- list()

for (participant_id in names(participant.data.s1)) {
  
  participant.df <- participant.data.s1[[participant_id]]
  
  session1.matrix <- matrix(participant.df$session1, ncol = 1)
  
  res.ecp <- e.cp3o_delta(Z = session1.matrix, K = 8, alpha = 1)
  
  pt.ecp <- get_earliest_cp(res.ecp$estimates, s)
  
  change.points.ecp.s1[[participant_id]] <- list(session1_cp = pt.ecp)
}

change.points.ecp.df.s1 <- do.call(rbind, lapply(change.points.ecp.s1, function(x) as.data.frame(t(unlist(x)))))
rownames(change.points.ecp.df.s1) <- names(change.points.ecp.s1)

save(change.points.ecp.df.s1, file = "./data/change.points.ecp.df.s1")
```

#### mcp
```{r include=FALSE, eval=FALSE}
change.points.mcp.s1 <- list()

options(mc.cores = 3)
prior <- list(
  cp_1 = "dunif(50, 150)"
)

model <- list(
  y ~ 1 + sigma(1),
    ~ 0 + x + sigma(1)
)

for (participant_id in names(participant.data.s1)) {
  
  participant.df <- participant.data.s1[[participant_id]]
  
  df.s1 <- tibble(x = s, y = participant.df$session1)
  
  fit1 <- mcp(model, prior = prior, data = df.s1, cores = 3, chains = 3)
  
  summary.fit.median1 <- summary.mcpfit.median(fit1)
  
  pt1.mcp <- round(summary.fit.median1[1,2],0)
  
  change.points.mcp.s1[[participant_id]] <- list(session1_mcp = pt1.mcp)
}

change.points.mcp.df.s1 <- do.call(rbind, lapply(change.points.mcp.s1, function(x) as.data.frame(t(unlist(x)))))
rownames(change.points.mcp.df.s1) <- names(change.points.mcp.s1)
save(change.points.mcp.df.s1, file = "./data/change.points.mcp.df.s1")
```

#### bcp
```{r}
change.points.bcp.s1 <- list()

for (participant_id in names(participant.data.s1)) {
  
  participant.df <- participant.data.s1[[participant_id]]
  
  df.s1 <- tibble(x = s, y = participant.df$session1)
  
  res <- bcp(df.s1$y, p0 = 0.1, mcmc = 3000)
  
  pt1.bcp <- s[which.max(res$posterior.prob)]
  
  change.points.bcp.s1[[participant_id]] <- list(session1_bcp = pt1.bcp)
}

change.points.bcp.df.s1 <- do.call(rbind, lapply(change.points.bcp.s1, function(x) as.data.frame(t(unlist(x)))))
rownames(change.points.bcp.df.s1) <- names(change.points.bcp.s1)
save(change.points.bcp.df.s1, file = "./data/change.points.bcp.df.s1")
```


#### summary statistics

```{r}

load(file = "./data/change.points.cp.df.s1")
load(file = "./data/change.points.pelt.df.s1")
load(file = "./data/change.points.ecp.df.s1")
load(file = "./data/change.points.mcp.df.s1")
load(file = "./data/change.points.bcp.df.s1")
# load(file = "./data/cs.data.s1")

# Medians
session1.120.median.cp <- round(median(change.points.cp.df.s1$session1.cp),0)

session1.120.median.pelt <- median(change.points.pelt.df.s1$session1.cp)

session1.120.median.ecp <- median(change.points.ecp.df.s1$session1_cp, na.rm = TRUE)

session1.120.median.mcp <- median(change.points.mcp.df.s1$session1_mcp)

session1.120.median.bcp <- median(change.points.bcp.df.s1$session1_bcp)
```


#### BinSeg distribution plot
```{r}
library(tidyr)

df <- change.points.cp.df.s1 |> pivot_longer(cols = everything(), names_to = "session", values_to = "change.points")

categ.palette <- c("#000000", "#E69F00", "#009E73", "#0072B2", "#D55E00", "#CC79A7")

p.s1.cp <- ggplot(data = df, aes(x = change.points, color = session)) + theme_gar +
  geom_freqpoly(binwidth = 1, size = .6, na.rm = TRUE, alpha = 0) +  
  scale_color_manual(values = categ.palette) +
  labs(x = "Time in ms", y = "Frequency", color = "Session") +
  ggtitle(paste("BinSeg, median onset =",session1.120.median.cp,"ms")) +
  geom_vline(xintercept = session1.120.median.cp, linetype = "dashed") +
  xlim(xa,xb) + ylim(0,11) + 
  geom_density(aes(y = ..scaled..*8.5), alpha = .3, adjust = .1, colour = "#E69F00", linewidth = .8) +
  theme(legend.position = "none", plot.title = element_text(size = 20))
p.s1.cp
```

#### PELT distribution plot
```{r}
df <- change.points.pelt.df.s1 |> pivot_longer(cols = everything(), names_to = "session", values_to = "change.points")

categ.palette <- c("#000000", "#E69F00", "#009E73", "#0072B2", "#D55E00", "#CC79A7")

p.s1.pelt <- ggplot(data = df, aes(x = change.points, color = session)) + theme_gar +
  geom_freqpoly(binwidth = 1, size = .6, na.rm = TRUE, alpha = 0) +  
  scale_color_manual(values = categ.palette) +
  labs(x = "Time in ms", y = "Frequency", color = "Session") +
  ggtitle(paste("PELT, median onset =",session1.120.median.pelt,"ms")) +
  geom_vline(xintercept = session1.120.median.pelt, linetype = "dashed") +
  xlim(xa,xb) + ylim(0,11) + 
  geom_density(aes(y = ..scaled..*8.5), alpha = .3, adjust = .1, colour = "#CC79A7", linewidth = .8) +
  theme(legend.position = "none", plot.title = element_text(size = 20)) 
p.s1.pelt
```

#### cp3o_delta distribution plot
```{r}
df <- change.points.ecp.df.s1 |> pivot_longer(cols = everything(), names_to = "session", values_to = "change.points")

categ.palette <- c("#000000", "#E69F00", "#009E73", "#0072B2", "#D55E00", "#CC79A7")

p.s1.ecp <- ggplot(data = df, aes(x = change.points, color = session)) + theme_gar +
  geom_freqpoly(binwidth = 1, size = .6, na.rm = TRUE, alpha = 0) +  
  scale_color_manual(values = categ.palette) +
  geom_density(aes(y = ..scaled..*8.5), alpha = .3, adjust = .1, colour = "#0072B2", linewidth = .8) +
  labs(x = "Time in ms", y = "Frequency", color = "Session") +
  geom_vline(xintercept = session1.120.median.ecp, linetype = "dashed") +
  ggtitle(paste("cp3o_delta, median onset =",session1.120.median.ecp,"ms")) +
  xlim(xa,xb) + ylim(0,11) + 
  theme(legend.position = "none", plot.title = element_text(size = 20))
p.s1.ecp
```

#### mcp distribution plots
```{r}
df <- change.points.mcp.df.s1 |> pivot_longer(cols = everything(), names_to = "session", values_to = "change.points")

categ.palette <- c("#000000", "#E69F00", "#009E73", "#0072B2", "#D55E00", "#CC79A7")

p.s1.mcp <- ggplot(data = df, aes(x = change.points, color = session)) + theme_gar +
  geom_freqpoly(binwidth = 1, size = .6, na.rm = TRUE, alpha = 0) +  
  scale_color_manual(values = categ.palette) +
  labs(x = "Time in ms", y = "Frequency", color = "Session") +
  ggtitle(paste("mcp, median onset =",session1.120.median.mcp,"ms")) +
  geom_vline(xintercept = session1.120.median.mcp, linetype = "dashed") +
  xlim(xa,xb) + ylim(0,11) + 
  geom_density(aes(y = ..scaled..*8.5), alpha = .3, adjust = .1, colour = "#D55E00", linewidth = .8) +
  theme(legend.position = "none", plot.title = element_text(size = 20))
p.s1.mcp
```

#### cluster sum distribution plot
```{r}
df <- cs.data.s1 |> pivot_longer(cols = everything(), names_to = "session", values_to = "change.points")

categ.palette <- c("#000000", "#E69F00", "#009E73", "#0072B2", "#D55E00", "#CC79A7")

session1.120.median.cs <- median(cs.data.s1$session1.cp)

p.s1.cs <- ggplot(data = df, aes(x = change.points, color = session)) + theme_gar +
  geom_freqpoly(binwidth = 1, size = .6, na.rm = TRUE, alpha = 0) +  
  scale_color_manual(values = categ.palette) +
  labs(x = "Time in ms", y = "Frequency", color = "Session") +
  ggtitle(paste("CS, median onset =",session1.120.median.mcp,"ms")) +
  geom_vline(xintercept = session1.120.median.cs, linetype = "dashed") +
  xlim(xa,xb) + ylim(0,11) + 
  geom_density(aes(y = ..scaled..*8.5), alpha = .3, adjust = .1, colour = "#009E73", linewidth = .8) +
  theme(legend.position = "none", plot.title = element_text(size = 20))
p.s1.cs
```

#### bcp distirbution plot
```{r}
df <- change.points.bcp.df.s1 |> pivot_longer(cols = everything(), names_to = "session", values_to = "change.points")

categ.palette <- c("#000000", "#E69F00", "#009E73", "#0072B2", "#D55E00", "#CC79A7")

p.s1.bcp <- ggplot(data = df, aes(x = change.points, color = session)) + theme_gar +
  geom_freqpoly(binwidth = 1, size = .6, na.rm = TRUE, alpha = 0) +  
  scale_color_manual(values = categ.palette) +
  labs(x = "Time in ms", y = "Frequency", color = "Session") +
  ggtitle(paste("bcp, median onset =",session1.120.median.bcp,"ms")) +
  geom_vline(xintercept = session1.120.median.bcp, linetype = "dashed") +
  xlim(xa,xb) + ylim(0,11) + 
  geom_density(aes(y = ..scaled..*8.5), alpha = .3, adjust = .1, colour = "#D55E00", linewidth = .8) +
  theme(legend.position = "none", plot.title = element_text(size = 20))
p.s1.bcp
```


#### joined plot
```{r fig.height=5, fig.width=10}
session1.120.median.cs <- median(cs.data.s1$session1.cp)
change.points.cp.df.s1 <- change.points.cp.df.s1  |>  mutate(method = "BinSeg")
change.points.pelt.df.s1 <- change.points.pelt.df.s1 |> mutate(method = "PELT")
change.points.ecp.df.s1 <- change.points.ecp.df.s1 |> mutate(method = "cp3o_delta")
change.points.mcp.df.s1 <- change.points.mcp.df.s1 |> mutate(method = "mcp")
cs.data.s1 <- cs.data.s1 |> mutate(method = "CS")
change.points.bcp.df.s1 <- change.points.bcp.df.s1 |> mutate(method = "bcp")

# Combine all data frames into one
combined_df <- bind_rows(change.points.cp.df.s1, change.points.pelt.df.s1, change.points.ecp.df.s1, change.points.mcp.df.s1, change.points.bcp.df.s1, cs.data.s1) |>
  pivot_longer(cols = starts_with("session"), names_to = "session", values_to = "change.points")

# Add median onset values for each method if needed for plotting
combined_df <- combined_df |>
  mutate(median_onset = case_when(
    method == "BinSeg" ~ session1.120.median.cp,
    method == "PELT" ~ session1.120.median.pelt,
    method == "cp3o_delta" ~ session1.120.median.ecp,
    method == "mcp" ~ session1.120.median.mcp,
    method == "CS" ~ session1.120.median.cs,
    method == "bcp" ~ session1.120.median.bcp
  ))

p.unified <- ggplot(combined_df, aes(x = change.points, color = method, group = method)) +
  geom_freqpoly(aes(y = ..count..), binwidth = 1, size = .6, alpha = 0) +  # Transparent frequency polygon
  geom_density(aes(y = ..scaled..*8.5), adjust = 0.1, size = 0.8, fill = NA, linetype = "solid") +  # Density plot
  scale_color_manual(values = c(
    "BinSeg" = "#E69F00", 
    "PELT" = "#D55E00", 
    "cp3o_delta" = "#009E73", 
    "mcp" = "#CC79A7", 
    "CS" = "#0072B2",
    "bcp" = "#000000"
  )) +
  labs(x = "Time in ms", y = "Frequency", color = "Method") + 
  theme_gar +
  # theme(legend.position = c(0.4, 0.9),  # Position inside the plot
  #       legend.justification = c("right", "top"),  # Anchors legend at position
  #       legend.box.background = element_rect(color = "black", size = 0.5),  # Optional border around legend
  #       legend.background = element_rect(fill = "white", size = 0.5, linetype = "solid")) +
  theme(legend.position = "bottom") +
  xlim(-400, 400) + ylim(0, 11) +
  geom_vline(xintercept = session1.120.median.cp, colour = "#E69F00", linetype = "dashed") +
  geom_vline(xintercept = session1.120.median.cs, colour = "#0072B2", linetype = "dashed") +
  geom_vline(xintercept = session1.120.median.pelt, colour = "#D55E00", linetype = "dashed") +
  geom_vline(xintercept = session1.120.median.mcp, colour = "#CC79A7", linetype = "dashed") +
  geom_vline(xintercept = session1.120.median.ecp, colour = "#009E73", linetype = "dashed") +
  geom_vline(xintercept = session1.120.median.bcp, colour = "#000000", linetype = "dashed")
p.unified

ggsave(filename = "./figures/real_data_120session.pdf", width = 8, height = 5)
```


```{r fig.height=8, fig.width=10}
cowplot::plot_grid(p.unified, p.s1.cp, p.s1.pelt, p.s1.ecp, p.s1.mcp, p.s1.cs,
                   nrow = 3,
                   labels = c("A", "B", "C", "D", "E", "F"),
                   label_size = 15)
```

Bieniek MM, Bennett PJ, Sekuler AB, Rousselet GA (2016) A robust and representative lower bound on object processing speed in humans. The European Journal of Neuroscience, 44(2), 18041814. https://doi.org/10.1111/ejn.13100